# An Empirical Study on the Execution Time and Performance of Python code using Python Energy Efficiency Guidelines -- NNDEV

## Setup (Installations)
```{r, results='hide', message=FALSE}
install.packages(c("tidyverse", "ggplot2", "bestNormalize", "ARTool", "lmerTest", "xtable"))
library(tidyverse)
library(ggplot2)
library(bestNormalize)
library(ARTool)
library(lmerTest)
library(xtable)
```

## 1. Data Loading and Inspection

```{r}
setwd("/Users/nhimai/Documents/MCs/Green-Lab/green-lab-project-nndev/data-analysis")
dat_data <- read.csv("../benchmarks/experiments/python_guidelines_20251015_152255/run_table.csv") %>%
  select(guideline, code, treatment, run_number, execution_time, execution_time, cpu_usage, memory_usage) %>%
  mutate(
    guideline = factor(guideline),
    code = factor(code),
    treatment = factor(treatment),
    run_number = factor(run_number)
  )

dat_data <- dat_data %>%
  filter(!is.na(execution_time) & !is.na(guideline) & !is.na(treatment))

glimpse(dat_data)
summary(dat_data)
head(dat_data)
```

## 2. Visual Exploration via Plotting

### 2.1. Histogram of execution_time

### 2.1.1. Overall

```{r}
ggplot(dat_data, aes(x = execution_time)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Histogram of Execution Time",
    x = "Execution Time",
    y = "Frequency"
  )
```

#### 2.1.2. Split by treatment
```{r}
ggplot(dat_data, aes(x = execution_time, fill = treatment)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  facet_wrap(~ guideline, scales = "free_x") +
  theme_minimal() +
  labs(
    title = "Execution Time Distribution by Treatment",
    x = "Execution Time",
    y = "Frequency",
    fill = "Treatment"
  )
```

### 2.2. Box plot of execution_time grouped by treatment and guideline
```{r}
ggplot(dat_data, aes(x = treatment, y = execution_time, fill=guideline)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Execution Time by Treatment",
    x = "Treatment",
    y = "Execution Time",
    fill = "Guideline"
  )
```

#### 2.3. Scatterplots of execution_time across trials: define "experiment" as guideline + treatment
```{r}
dat_data <- dat_data %>%
  mutate(
    experiment = interaction(guideline, treatment),
    run_number = factor(run_number, levels = paste0("r", 1:15), ordered = TRUE)
    )

ggplot(dat_data, aes(x = run_number, y = execution_time, color = experiment)) +
  geom_point(alpha = 0.8, size = 2) +
  theme_minimal() +
  labs(
    title = "Execution Time Across Runs",
    x = "Run Number",
    y = "Execution Time",
    color = "Experiment"
  )
```


## 3. Data Normality and Normalization

### 3.1. Shapiro-Wilk test
```{r}
# Run Shapiro-Wilk for each group
normality_results <- dat_data %>%
  group_by(guideline, treatment) %>%
  summarise(
    shapiro_p = shapiro.test(execution_time)$p.value,
    .groups = "drop"
  )

normality_results
```

- Interpretation: if p < 0.05, reject normality (data is not normal).

### 3.2. Normalize the execution_time column
```{r}
bn <- bestNormalize(dat_data$execution_time)
dat_data$norm_execution_time <- bn$x.t
```

### 3.3. Visualization of normalized data and repeat normality tests

#### 3.3.1. Histogram of normalized data
```{r}
ggplot(dat_data, aes(x = norm_execution_time)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  theme_minimal() +
  labs(
    title = "Histogram of Normalized Execution Time",
    x = "Normalized Execution Time",
    y = "Frequency"
  )
```

#### 3.3.2. Shapiro-Wilk again for normalized values
```{r}
normality_results_norm <- dat_data %>%
  group_by(guideline, treatment) %>%
  summarise(
    shapiro_p = shapiro.test(norm_execution_time)$p.value,
    .groups = "drop"
  )

normality_results_norm
```
- If most *shapiro_p* values after normalization are greater than 0.05, then distributions can be considered normal, and parametric tests (t-test) are justified.
- If many are still below 0.05, non-parametric test (Wilcoxon) will be the case.

## 4. Non-parametric test

### 4.1. Wilcoxon’s test

#### Compare entire set
```{r, results='hide', message=FALSE}
all_no_guideline_j = dat_data[dat_data$treatment == 'no_guideline',]$execution_time
all_guideline_j = dat_data[dat_data$treatment == 'guideline',]$execution_time
wilcox.test(all_no_guideline_j, all_guideline_j)

# g1_no_guideline = dat_data[dat_data$treatment == 'no_guideline' & dat_data$guideline == 'G1',]$execution_time
# g1_guideline = dat_data[dat_data$treatment == 'guideline' & dat_data$guideline == 'G1',]$execution_time
# wilcox.test(g1_no_guideline, g1_guideline)
```

#### Compare different treatments in terms of guidelines
```{r}
guidelines <- unique(dat_data$guideline)
wilcox_results <- map_dfr(guidelines, function(g) {
  subset_data <- dat_data %>% filter(guideline == g)
  
  no_guideline_j <- subset_data %>% filter(treatment == "no_guideline") %>% pull(execution_time)
  guideline_j     <- subset_data %>% filter(treatment == "guideline") %>% pull(execution_time)
  
  test_result <- wilcox.test(no_guideline_j, guideline_j)
  
  tibble(
    guideline = g,
    p_value = test_result$p.value,
    statistic = test_result$statistic,
    method = test_result$method
  )
})

print(wilcox_results)
```

### 4.2. ART (Aligned Rank Transform) model
```{r}
art_model <- art(execution_time ~ guideline * treatment + (1|run_number), data = dat_data)
summary(art_model)
```

### 4.3. LMER model
```{r, results='hide', message=FALSE}
lmer_model = lmer(execution_time ~ guideline * treatment + (1|run_number), data = dat_data)
summary(lmer_model)
anova(lmer_model)
```

### 4.4. Check residuals for normality
```{r}
res_model <- resid(model)
qqnorm(res_model)
qqline(res_model, col = "red", lwd = 2)
shapiro.test(res_model)
```

## 5. Summary plot

## 5.1. Boxplot of execution_time by code, faceted by guideline
```{r}
ggplot(dat_data, aes(x = as.factor(code), y = execution_time)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  facet_wrap(~ guideline, scales = "free") +
  labs(x = "Code", y = "Execution Time", title = "Execution Time Distribution by Code") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 5.3. Scatter plot of execution_time over run_number, colored by guideline
```{r}
ggplot(dat_data, aes(x = run_number, y = execution_time, color = guideline)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(x = "Run Number", y = "Execution Time", title = "Execution Time over Run Number") +
  theme_minimal() +
  theme(legend.position = "right")
```