# An Empirical Study on the Energy Usage and Performance of Python codes using Python Energy Efficiency Guidelines -- NNDEV

## Setup
```{r, results='hide', message=FALSE}
install.packages(c("tidyverse", "ggplot2", "bestNormalize", "ARTool", "lmerTest", "xtable"))
library(tidyverse)
library(ggplot2)
library(bestNormalize)
library(ARTool)
library(lmerTest)
library(xtable)
```

## 1. Data Loading and Inspection

### 1.1. Load data
```{r}
setwd("/Users/ntd432/Desktop/SE4GD/VU/GreenLab/project/green-lab-project-nndev/data-analysis")
dat_data <- read.csv("raw-data/dummy_data.csv") %>%
  select(guideline, code, treatment, run_number, energy_usage, execution_time, cpu_usage, memory_usage) %>%
  mutate(
    guideline = factor(guideline),
    code = factor(code),
    treatment = factor(treatment),
    run_number = factor(run_number)
  )

glimpse(dat_data)
summary(dat_data)
head(dat_data)
```

### 1.2. Histogram overall

```{r}
# Histogram of overall energy usage
ggplot(dat_data, aes(x = energy_usage)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Histogram of Energy Usage",
    x = "Energy Usage",
    y = "Frequency"
  )
```

#### 1.2.1. Split by treatment
```{r}
# Histogram split by treatment
ggplot(dat_data, aes(x = energy_usage, fill = treatment)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  facet_wrap(~ guideline, scales = "free_x") +
  theme_minimal() +
  labs(
    title = "Energy Usage Distribution by Treatment",
    x = "Energy Usage",
    y = "Frequency",
    fill = "Treatment"
  )
```

#### 1.2.2. Split by run
```{r}
# Define "experiment" as library + dataframe_size
dat_data <- dat_data %>%
  mutate(
    experiment = interaction(guideline, treatment),
    run_number = factor(run_number, levels = paste0("r", 1:15), ordered = TRUE)
    )

ggplot(dat_data, aes(x = run_number, y = energy_usage, color = experiment)) +
  geom_point(alpha = 0.8, size = 2) +
  theme_minimal() +
  labs(
    title = "Energy Usage Across Runs",
    x = "Run Number",
    y = "Energy Usage",
    color = "Experiment"
  )
```

### 1.3. Box plot
```{r}
ggplot(dat_data, aes(x = treatment, y = energy_usage, fill=guideline)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Energy Usage by Treatment",
    x = "Treatment",
    y = "Energy Usage",
    fill = "Guideline"
  )
```

## 2. Data Normality and Normalization

### 2.1. Q-Q plot
### 2.2. Shapiro-Wilk test
```{r}
# Run Shapiro-Wilk for each group
normality_results <- dat_data %>%
  group_by(guideline, treatment) %>%
  summarise(
    shapiro_p = shapiro.test(energy_usage)$p.value,
    .groups = "drop"
  )

normality_results
```
Interpretation: if p < 0.05, reject normality (data is not normal).

### 2.3. Normalize data
```{r}
bn <- bestNormalize(dat_data$energy_usage)
dat_data$norm_energy_usage <- bn$x.t
```

### 2.4. Visualization of normalized data (maybe do the test again)
```{r}
# Histogram of normalized values
ggplot(dat_data, aes(x = norm_energy_usage)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  theme_minimal() +
  labs(
    title = "Histogram of Normalized Energy Usage",
    x = "Normalized Energy Usage",
    y = "Frequency"
  )

# Shapiro-Wilk again for normalized values
normality_results_norm <- dat_data %>%
  group_by(guideline, treatment) %>%
  summarise(
    shapiro_p = shapiro.test(norm_energy_usage)$p.value,
    .groups = "drop"
  )

normality_results_norm
```

- If most *shapiro_p* values after normalization are greater than 0.05, then distributions can be considered normal, and parametric tests (t-test) are justified.
- If many are still below 0.05, non-parametric tests (Wilcoxon, ART).

## 3. RQ1 tests (TBD)

## 4. RQ2 tests (TBD)

## 5. Summary plot
